---
title: "GEM Gas Exploratory Analysis"
output: 
  pdf_document:
    highlight: tango
---

```{r include=F}
library(tidyverse)
library(sf)
library(tmap)
```

```{r include=F}
gas_unit <- read_csv("../data/raw_data/gas_unit.csv")
gas_unit <- 
  gas_unit %>% mutate(Capacity = na_if(`Capacity elec. (MW)`, "not found") 
                      %>% as.numeric()) %>% 
  mutate(StartYear = na_if(`Start year`, "not found") 
                      %>% as.numeric()) 
```
```{r include=F}

China_gas <- gas_unit %>% filter(Country == "China")
operating_gas <- China_gas %>% filter(Status == "operating")
operating_gas %>% pull(Capacity) %>% sum(na.rm=T) 
operating_gas %>% pull(StartYear) %>% range(na.rm=T)
```

## Descriptive Statistics

**Total Number of plants:** There are a total of 680 gas units in GEM gas 
plants tracker.

**Time Coverage:** Start years of the plants ranges from 1991 to 2022

**Operating Plants and total capacity: **
Among these plants, 349 of them are operating, with a total installed capacity
of 103480 MW. 
This number is very close to the number published by CEC's published number 
for 2021, which is 107040 MW. 

CEC's number for 2021
https://cec.org.cn/detail/index.html?3-304834



