---
title: "GEM China Coal Plants Retirement Visualization Exercise"
author: Chi Gao
date: April 14, 2022
output:
  prettydoc::html_pretty:
    theme: leonids 
---

```{r include=F}
knitr::opts_chunk$set(echo=F, warning=F, message=F)
```

```{r include=F}
library(tidyverse)
library(plotly)
library(prettydoc)

short_plot <- readRDS("../temp/short_plot.rds")
mid_plot <- readRDS("../temp/mid_plot.rds")
long_plot <- readRDS("../temp/long_plot.rds")
retirement_over_time_plot <- readRDS("../temp/retirement_over_time_plot.rds")
lifespan_plot <- readRDS("../temp/lifespan_plot.rds")
```

# Introduction 

The main purpose of this visualization exercise is to illustrate China's coal plants 
natural retirement patterns. That is, assuming China is not constructing any
more new coal-fired power plants and all coal plants were to live up to its full planned 
lifetime, what does China's coal-fired power generation capacity trajectory look like? 

All the the plots here are made using GEM's [Coal Power Plants Tracker](https://globalenergymonitor.org/projects/global-coal-plant-tracker/download-data/).

# Historical Retirement patterns

To estimate the trajectory of coal generation capacity, it is imperative to inspect 
the historical 
retirement records. To this end, I first select only the retired plants and plot out the 
relationship between plant operation start time and plant lifetime. 

```{r}
lifespan_plot <- 
lifespan_plot + geom_hline(yintercept = 22, linetype="dashed", color="red") + 
  geom_text(x=1960, y=20, label="Average Lifetime: 22", color="red") 

lifespan_plot %>% ggplotly()
```

Curiously, there is a significant linear relationship between operation start year and 
the lifespan of coal-fired power plants. This strong association can be attributed to China's 
slew of campaigns targeting at small coal plants to address overcapacity problems. One 
regulatory document issued in 2005 that heralds these retirement and restructuring 
campaigns is the 
[Decision of the State Council on Promulgating the “Interim Provisions on Promoting Industrial Structure Adjustment” for Implementation](http://www.lawinfochina.com/display.aspx?lib=law&id=4828), which highlighted China's pressing needs to "build large-type coal bases, adjust
and transform medium and small coal mines, firmly eliminate the small coal mines
failing to meet work safety conditions and wasting or destroying resources."
To see how these policies are affecting the retirement of coal-fired power plants, I plotted
the histogram of coal plant retirement over time: 

```{r}
retirement_over_time_plot <- 
  retirement_over_time_plot + theme(text=element_text(size=11)) + 
  scale_x_continuous(breaks=seq(1990, 2020, 5)) + 
  labs(y = "Count", x = "Retirement start year")

retirement_over_time_plot %>% ggplotly()
```

Indeed, starting from 2005, there is a sudden increase in the amount of retired 
coal plants; many smaller plants were shut down before their planned retirement are due. 

# Natural coal plants phaseout trajectory

This section aims to summarize the coal generation capacity trajectory under the natural 
phaseout assumption. That is, no new coal plants will be constructed unless the plant is 
currently under construction or already acquired permits. These sets of graphs are made 
to gauge how much early retirement and retrofitting efforts are required to achieve 
carbon neutrality under different planned lifetime schemes. 

```{r}
short_plot <- 
  short_plot + theme(text=element_text(size=11)) + 
  scale_x_continuous(breaks=seq(2000, 2080, 10), limits = c(2000, 2080)) + 
  labs(x = "Year")
ggplotly(short_plot)
```

```{r}
mid_plot <- 
  mid_plot + theme(text=element_text(size=11)) +
  scale_x_continuous(breaks=seq(2000, 2080, 10), limits = c(2000, 2080)) + 
  labs(x = "Year")

ggplotly(mid_plot)
```

```{r}
long_plot <- 
  long_plot + theme(text=element_text(size=11)) + 
  scale_x_continuous(breaks=seq(2000, 2080, 10), limits = c(2000, 2080)) + 
  labs(x = "Year")

ggplotly(long_plot)
```

As we can see, in the more realistic lifetime scenarios of 30 years and 40 years, natural 
phaseout is not satisfactory per se. In my future work, I will identify retrofitting and 
retirement capacity of each province. 

Note: In part to address some missing data, some assumptions are made to produce these three plots above.

1. All plants under construction will complete construction, assuming it will 
take five years. 
1. Planned lifetime is only applied to plants without known planned retirement date. 
1. For coal plants that are permitted yet haven't started their construction, a 5-10 years 
pre-construction period is assigned randomly before entering the construction phase. 
1. Given the strong linear relationship explored in previous section, lifetime of those plants
whose retirement year is unknown is predicted using their operation start year. 
1. Cancelled, shelved, pre-permit, mothballed, and announced projects, as well as 
operating plants with no starting  and retirement years are all not considered in the 
visualization exercises. 



